Palomar 5 Gaps

### 2024-10-11 paper/main.tex
rewrote the edit

  \paola{Postulated about a century ago, to explain the large velocity dispersions of galaxies inside clusters \citep[see][]{zwicky33}\footnote{The English translation of the original paper is available at \citet{zwicky_EN}} and the rotation curves of disk galaxies as well \citep{freeman70, rubin70}, different  astronomical measurements conducted over the last century point to the existence of dark matter, from temperature fluctuations in the Cosmic Microwave Background  to weak and strong gravitational lensing \citep[see, among many, ][]{planck20, mandelbaum06, hsueh20}. Dark matter constitutes the dominant ($\sim$~85\%) contributor to matter in the Universe \citep{planck20}, and shapes the building-up of its structures over time \citep{padmanabhan93, bertschinger98, springel05}.}

>>>>

  % \paola{Postulated about a century ago, to explain the large velocity dispersions of galaxies inside clusters \citep[see][]{zwicky33}\footnote{The English translation of the original paper is available at \citet{zwicky_EN}} and the rotation curves of disk galaxies as well \citep{freeman70, rubin70}, different  astronomical measurements conducted over the last century point to the existence of dark matter, from temperature fluctuations in the Cosmic Microwave Background  to weak and strong gravitational lensing \citep[see, among many, ][]{planck20, mandelbaum06, hsueh20}. Dark matter constitutes the dominant ($\sim$~85\%) contributor to matter in the Universe \citep{planck20}, and shapes the building-up of its structures over time \citep{padmanabhan93, bertschinger98, springel05}.}

  Over a century ago \citet{zwicky33}\footnote{The English translation of the original paper is available at \citet{zwicky_EN}} postulated the existence of dark matter to explain the excessively large velocity dispersions of galaxies within their galaxy clusters. Since this time, mounting evidence has been assembled to support this claim. \citet{freeman70} and \citet{rubin70} provided some of the first measurements of the rotation curves of disk galaxies that are too fast moving to be driven by ordinary matter alone. The temperature fluctuations in the Cosmic Microwave Background radation are well explained by the presence of dark matter, which is well studied and documented by the Planck Mission, a review of which is presented by \citet{planck20}. The location and mass of dark matter on the large scale can be inferred and spatially mapped through strong and weak gravitational lensing, as demonstrated in \citet{planck20, mandelbaum06, hsueh20}. Dark matter contributes $\sim$~85~\% of the matter in the universe \citep{planck20} and is crucial for driving the hierarchical growth of ordinary in the universe. See \citet{padmanabhan93} for XXX, \citet{bertschinger98} for XXX and \citet{springel05} for XXX.
